\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-12 tex/paper.tex
add figure

\usepackage{bm}
\usepackage{cancel}
\usepackage{bbold}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{color}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

In this work, we repurpose adversarial training as a means to constraint the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
particular, we pit $f$ against an adversary model $r := p_{\theta_r}(z |
f(X;\theta_f))$ of parameters $\theta_r$ and associated loss ${\cal
L}_r(\theta_f, \theta_r)$. This model takes  as input realizations of $f(X;
\theta_f)$, for the current value $\theta_f$ of $f$ parameters, and produces as
output a function $p_{\theta_r}(z | f(X;\theta_f))$ modeling the posterior
probability density that $z$ parameterizes the sample $X$ observed through
$f(\cdot;\theta_f)$. If $Z$ takes discrete values, then $p_{\theta_r}$ can be
represented as a probabilistic classifier $\mathbb{R} \mapsto \mathbb{R}^{|{\cal
Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$) is the estimated
probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly, if $Z$ takes
continuous values and if we assume some parameteric distribution for
$Z|f(X;\theta_f)$ (e.g., a gaussian), then $p_{\theta_r}$ can be represented as
network whose output $j$ is the estimated value of the corresponding parameter
of that distribution (e.g., its mean and variance). As in
\citep{nix1994estimating}, the estimated probability density
$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$
under that distribution.  Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
then the corresponding correlation can be captured by $r$. By contrast, if
$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
perform poorly and be close to random guessing. Training $f$ such that it
additionally minimizes the performance of $r$ therefore acts as a regularization
towards Eqn.~\ref{eqn:criterion}.

@@

\glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}
\glnote{Add summary figure.}

>>>>

\usepackage{bm}
\usepackage{cancel}
\usepackage{bbold}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{color}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}

@@

\begin{figure*}
    \usetikzlibrary{arrows}
    \def\layersep{1cm}

    \begin{tikzpicture}[shorten >= 1pt, ->, node distance=\layersep,scale=.8]
    \tikzstyle{neuron} = [circle, minimum size=0.25cm, draw=black!30, line width=0.3mm, fill=white]

    % Classifier f
    \node at (2,0) {Classifier $f$};
    \draw (-1,-0.5) rectangle (4,-5.5);

    \path[->, shorten >= 0pt] (-2,-3) edge (-1,-3);
    \node[left] at (-2,-3) {$X$};

    \path[-o, shorten >= 0pt] (1.5,-6.5) edge (1.5,-5.5);
    \node[below] at (1.5,-6.5) {$\theta_f$};

    \path[->, shorten >= 0pt] (3.5,-3) edge (6.5,-3);
    \node[above] at (5.25,-3) {$f(X;\theta_f)$};

    \foreach \name / \y in {1,...,3}
        \node[neuron] (f-I-\name) at (-0.5,-1-\y) {};

    \foreach \name / \y in {1,...,5}
        \node[neuron] (f-H1-\name) at (-0.5cm+\layersep,-\y cm) {};
    \foreach \name / \y in {1,...,5}
        \node[neuron] (f-H2-\name) at (-0.5cm+3*\layersep,-\y cm) {};

    \node[neuron] (f-O) at (-0.5cm+4*\layersep,-3cm) {};

    \foreach \source in {1,...,3}
        \foreach \dest in {1,...,5}
            \path[black!20] (f-I-\source) edge (f-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black!30] (f-H2-\source) edge (f-O);

    \node[black!20] at (1.5,-3) {...};

    % Adversary r
    \node at (9,0) {Adversary $r$};
    \draw (6.5,-0.5) rectangle (11.5,-5.5);

    \node[above] at (12.75,-2.5) {$\mu(f(X;\theta_f))$};
    \path[-o, shorten >= 0pt] (11,-2.5) edge (14,-2.5);
    \node[above] at (12.75,-3.5) {$\sigma(f(X;\theta_f))$};
    \path[-o, shorten >= 0pt] (11,-3.5) edge (14,-3.5);

    \path[-o, shorten >= 0pt] (9,-6.5) edge (9,-5.5);
    \node[below] at (9,-6.5) {$\theta_r$};

    \foreach \name / \y in {1,...,1}
        \node[neuron] (r-I-\name) at (7,-2-\y) {};

    \foreach \name / \y in {1,...,5}
        \node[neuron] (r-H1-\name) at (7cm+\layersep,-\y cm) {};
    \foreach \name / \y in {1,...,5}
        \node[neuron] (r-H2-\name) at (7cm+3*\layersep,-\y cm) {};

    \node[neuron] (r-O1) at (7cm+4*\layersep,-2.5cm) {};
    \node[neuron] (r-O2) at (7cm+4*\layersep,-3.5cm) {};

    \foreach \source in {1,...,1}
        \foreach \dest in {1,...,5}
            \path[black!20] (r-I-\source) edge (r-H1-\dest);

    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O1);
    \foreach \source in {1,...,5}
        \path[black!30] (r-H2-\source) edge (r-O2);

    \node[black!20] at (9,-3) {...};

    \draw (14,-1.5) rectangle (17,-4.5);
    \path[->, shorten >= 0pt] (15.5,-0.5) edge (15.5,-1.5);
    \node[above] at (15.5,-0.5) {$z$};
    \path[->, shorten >= 0pt] (15.5,-4.5) edge (15.5,-5.5);
    \node[below] at (15.5,-5.5) {$p_{\theta_r}(z|f(X;\theta_f))$};
    \node at (15.5,-3) {Truncated ${\cal N}$};

    \end{tikzpicture}

    \caption{Architecture for the adversarial training of a binary classifier $f$ against a continuous parameter $Z$.}
    \label{fig:architecture}
\end{figure*}

In this work, we repurpose adversarial training as a means to constraint the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. As
illustrated in Figure~\ref{fig:architecture}, we pit $f$ against an adversary
model $r := p_{\theta_r}(z | f(X;\theta_f))$ of parameters $\theta_r$ and
associated loss ${\cal L}_r(\theta_f, \theta_r)$. This model takes  as input
realizations of $f(X; \theta_f)$, for the current value $\theta_f$ of $f$
parameters, and produces as output a function $p_{\theta_r}(z | f(X;\theta_f))$
modeling the posterior probability density that $z$ parameterizes the sample $X$
observed through $f(\cdot;\theta_f)$. If $Z$ takes discrete values, then
$p_{\theta_r}$ can be represented as a probabilistic classifier $\mathbb{R}
\mapsto \mathbb{R}^{|{\cal Z|}}$ whose output $j$ (for $j=1, \dots, |{\cal Z}|$)
is the estimated probability mass $p_{\theta_r}(z_j|f(X;\theta_f))$. Similarly,
if $Z$ takes continuous values and if we assume some parameteric distribution
for $Z|f(X;\theta_f)$ (e.g., a truncated gaussian over a bounded support), then $p_{\theta_r}$ can be represented
as network whose output $j$ is the estimated value of the corresponding
parameter of that distribution (e.g., its mean and variance). As in
\citep{nix1994estimating}, the estimated probability density
$p_{\theta_r}(z|f(X;\theta_f))$ can then be evaluated for any $z \in {\cal Z}$
under that distribution.  Intuitively, if $p(f(X; \theta_f)|z)$ varies with $z$,
then the corresponding correlation can be captured by $r$. By contrast, if
$p(f(X; \theta_f)|z)$ is invariant with $z$, as we require, then $r$ should
perform poorly and be close to random guessing. Training $f$ such that it
additionally minimizes the performance of $r$ therefore acts as a regularization
towards Eqn.~\ref{eqn:criterion}.

@@

\glnote{Discuss that we could equivalently require an intermediate representation of the data (as learned internally by $f$) to be pivotal, instead of requiring $f(X)$ to be pivotal.}
